---
title: "ParseData"
output: html_document
date: "2024-09-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo =FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(stringr)
library(scales)
library(viridis)
library(mnormt)
library(ggfortify)
library(survival)
library(survminer)
library(Rcpp)

expit <- function(x){
  to_ret <- exp(x) / (1+exp(x))
  if (is.na(to_ret)){return(1)}
  return(to_ret)
}

logit <- function(x){
  return(log(x/(1-x)))
}

Param2TranHelper <- function(p12,p21){
  tran <- matrix(0,2,2)
  tran[1,2] <- expit(p12)
  tran[1,1] <- 1- tran[1,2]
  tran[2,1] <- expit(p21)
  tran[2,2] <- 1 - tran[2,1]
  return(tran)
}

CalcPiHelper <- function(nu, nu2,age){
  pi_ind <- exp(nu*age + nu2*age^2)/sum(exp(nu*age + nu2*age^2))
  if (any(is.na(pi_ind))){
    pi_ind[is.na(pi_ind)] <- 1
    pi_ind <- pi_ind/sum(pi_ind)
  }
  return(pi_ind)
}

CalcPi <- function(nu,nu2,age_vec){
  return(t(sapply(age_vec/10,CalcPiHelper,nu=nu, nu2=nu2)))
}

Params2TranVectorTresid <- function(re_ind,len,params_tran){
  return(t(sapply(c(1:(len)),FUN = Params2Tran,params_tran = params_tran,re_ind=re_ind)))
}

Params2Tran <- function(params_tran,time,re_ind){
  
  param_matrix <- matrix(params_tran[re_ind,],ncol=3,nrow=2, byrow = T)
  tran <- Param2TranHelper(param_matrix[1,1]+param_matrix[1,2]*cos(2*pi*time/96)+param_matrix[1,3]*sin(2*pi*time/96),
                           param_matrix[2,1]+param_matrix[2,2]*cos(2*pi*time/96)+param_matrix[2,3]*sin(2*pi*time/96))


  return(tran)
}

ParamsArray2DF <- function(params_tran_array){
  vcovar_num <- 2
  mix_num <- dim(params_tran_array)[1]
  tran_df <- data.frame(prob = c(),
                        type = c(),
                        time = c(),
                        age = c(),
                        weekend = c(),
                        mixture = c())
  for (re_ind in 1:mix_num){
    for (vcovar_ind in 1:vcovar_num){
      tran_mat <- Params2TranVectorTresid(re_ind,96,params_tran_array[,,vcovar_ind])
      tosleep <- tran_mat[,3]
      towake <- tran_mat[,2]
      
      tran_df_working <- data.frame(prob = c(tosleep,towake),
                                    type = rep(c("Falling Asleep", "Waking"),each= 96),
                                    time = rep(c(1:96)/4,2),
                                    weekend = vcovar_ind,
                                    mixture = re_ind)
      tran_df <- rbind(tran_df,tran_df_working)
    }
  }
  
  return(tran_df)
}

Q95 <- function(x){return(quantile(x,probs = c(.95)))}
Q5 <- function(x){return(quantile(x,probs = c(.05)))}

FirstDay2WeekInd <- function(first_day){
  weekday <- numeric(96)
  friday <- c(rep(0,68),rep(1,28))
  saturday <- numeric(96)+1
  sunday <- c(rep(1,68),rep(0,28))
  
  if (first_day == 1){
    covar_vec <- c(sunday,rep(weekday,4),friday,saturday,sunday,weekday)
  } else if (first_day == 2) {
    covar_vec <- c(rep(weekday,4),friday,saturday,sunday,rep(weekday,2))
  } else if (first_day == 3) {
    covar_vec <- c(rep(weekday,3),friday,saturday,sunday,rep(weekday,3))
  } else if (first_day == 4) {
    covar_vec <- c(rep(weekday,2),friday,saturday,sunday,rep(weekday,4))
  } else if (first_day == 5) {
    covar_vec <- c(weekday,friday,saturday,sunday,rep(weekday,4),friday)
  } else if (first_day == 6) {
    covar_vec <- c(friday,saturday,sunday,rep(weekday,4),friday,saturday)
  } else if (first_day == 7) {
    covar_vec <- c(saturday,sunday,rep(weekday,4),friday,saturday,sunday)
  }
  
  return(covar_vec)
}

setwd("../Data/")
load("NHANES_2011_2012_2013_2014.rda")
nhanes1 <- NHANES_mort_list[[1]] %>% filter(eligstat == 1)
nhanes2 <- NHANES_mort_list[[2]] %>% filter(eligstat == 1)
lmf_data <- rbind(nhanes1,nhanes2)


load("Wavedata_G.rda")
load("Wavedata_H.rda")
setwd("../Output/")

sim_num <- as.numeric(Sys.getenv('SLURM_ARRAY_TASK_ID'))

act_G <- wave_data_G[[1]]
act_H <- wave_data_H[[1]]
act <- rbind(act_G,act_H)
act <- t(act[,-1])
lod_act <- min(act[act>0],na.rm = T)
act0 <- act == 0
act <- log(act)
act[act0] <- log(lod_act * .99)

light_G <- wave_data_G[[2]]
light_H <- wave_data_H[[2]]
light <- rbind(light_G,light_H)
light <- t(light[,-1])
lod_light <- min(light[light>0],na.rm = T)
light0 <- light == 0
light <- log(light)
light[light0] <- log(lod_light * .99)

id_G <- wave_data_G[[3]]
id_H <- wave_data_H[[3]]
id <- rbind(id_G,id_H)

seqn_com_id <- id$SEQN %in% lmf_data$seqn
seqn_com_lmf <- lmf_data$seqn %in% id$SEQN

id <- id[seqn_com_id,]
act <- act[,seqn_com_id]
light <- light[,seqn_com_id]

lmf_data <- lmf_data[seqn_com_lmf,]

if (sum(id$SEQN - lmf_data$seqn) != 0){print("LMF NOT LINKED CORRECTLY")}

log_sweights_vec <- log(id$sweights/2)

id <- id %>% mutate(age_disc = case_when(age <=30 ~ 1,
                                         age <=50 & age > 30 ~ 2,
                                         age <=65 & age > 50 ~ 3,
                                         age > 65 ~ 4))

fcovar_num <- max(id$age_disc,na.rm = T)

# id <- id %>% mutate(age_disc = case_when(age <=30 ~ 1,
#                                          age <=40 & age > 30 ~ 2,
#                                          age <=50 & age > 40 ~ 3,
#                                          age <=60 & age > 50 ~ 4,
#                                          age <=70 & age > 60 ~ 5,
#                                          age > 70 ~ 6))

id <- id %>% mutate(pov_disc = floor(poverty)+1)

id <- id %>% mutate(bmi_disc = case_when(BMI <= 18.5 ~ 1,
                                         BMI <=25 & BMI > 18.5 ~ 2,
                                         BMI <=30 & BMI > 25 ~ 3,
                                         BMI <=35 & BMI > 30 ~ 4,
                                         BMI <=40 & BMI > 35 ~ 5,
                                         BMI > 40 ~ 6))

surv_event <- lmf_data$mortstat
surv_time <- lmf_data$permth_exm



#need to initialize starting values

# num_of_people <- 2000
# act <- act[,1:num_of_people]
# light <- light[,1:num_of_people]
# surv_event <- surv_event[1:num_of_people]
# surv_time <- surv_time[1:num_of_people]

first_day_vec <- as.numeric(id$PAXDAYWM)

vcovar_mat <- sapply(first_day_vec,FirstDay2WeekInd)
fcovar_vec <- id$age_disc

to_keep_inds <- !is.na(fcovar_vec)

act <- act[,to_keep_inds]
light <- light[,to_keep_inds]
vcovar_mat <- vcovar_mat[,to_keep_inds]
fcovar_vec <- fcovar_vec[to_keep_inds]
surv_event <- surv_event[to_keep_inds]
surv_time <- surv_time[to_keep_inds]


day_length <- dim(act)[1]
num_of_people <- dim(act)[2]

age_vec <-id$age

```
 
## BIC for each MHMM by size

7 mixtures has the lowest BIC, so we will analyze the results for 7 mixtures
 
```{r,echo=F}
file_list <- list.files()
file_list <- file_list[-length(file_list)]
num_files <- length(file_list)

bic_vec <- c()
for (i in 1:num_files){
  load(file_list[i])
  bic_vec <- c(bic_vec,to_save[[3]])
}

bic_df <- data.frame(BIC = bic_vec,
                     Mixture = c(2:(num_files+1)))

ymin <- .995 * min(bic_vec)
ymax <- 1.005 * max(bic_vec)

ggplot(bic_df, aes(x=Mixture, y=BIC)) +
  geom_bar(stat="identity") + 
  coord_cartesian(ylim=c(ymin,ymax)) + 
  ggtitle("BIC by Number of Mixtures")


load(file_list[which.min(bic_vec)])
# load(file_list[2])


mix_num <- dim(to_save[[1]][[1]])[1] 

init <- to_save[[2]][[1]]
params_tran_array <- to_save[[2]][[2]]
emit_act <- to_save[[2]][[3]]
emit_light <- to_save[[2]][[4]]
corr_mat <- to_save[[2]][[5]]
nu <- to_save[[2]][[6]]
nu2 <- to_save[[2]][[7]]
beta_vec <- to_save[[2]][[8]]
beta_age <- to_save[[2]][[9]]

# tran_df <- ParamsArray2DF(params_tran_array)
tran_df <- to_save[[2]][[10]]
re_prob <- to_save[[2]][[11]]
# weights_array_list <- to_save[[2]][[12]]

```



# Survival

```{r}
pi_l <- CalcPi(nu,nu2,age_vec)
print("The breakdown by cluster probability is as follows:")
round(colSums(re_prob)/sum(colSums(re_prob)),5)
print("")

print("Expected number of people in each cluster")
round(colSums(re_prob))
print("")

print("The Cox coefficient associated with an increase in one year of age is as follows:")
beta_age
print("")

print("The Cox coefficient associated with being in each cluster, with the first cluster as reference is as follows:")
beta_vec
print("")

mix_assignment <- apply(re_prob,1,which.max)

surv_data <- data.frame(time = rep(surv_time,each = mix_num),
                        status = rep(surv_event, each = mix_num),
                        group = rep(c(1:mix_num),length(surv_time)),
                        age = rep(age_vec,each = mix_num),
                        weights = as.vector(t(re_prob)))

surv_data$group <- as.factor(surv_data$group)
surv_data <- surv_data %>% filter(weights > 0)

fit <- survfit(Surv(time, status) ~ group, data = surv_data,weights = weights)

```

#### Survival curves by mixture assignment 

```{r}

autoplot(fit, conf.int = F,
         censor.shape = '.', censor.size = 10, facets = F, ncol = 2)+
    scale_color_viridis(discrete=TRUE,end = .95) 
```

### Log rank test

```{r}
survdiff(Surv(time, status) ~ group + age, data = surv_data, rho = 0)

```

### Cox CI

```{r}
coxph(Surv(time, status) ~ group + age, data = surv_data,weights = weights)
```

# Mixture Probabilities 


```{r}
pi_df <- melt(CalcPi(nu,nu2,c(18:80)))
colnames(pi_df) <- c("Age", "Mixture", "Prob")
pi_df$Age <- pi_df$Age + 17
pi_df$Mixture <- as.factor(pi_df$Mixture)

ggplot(pi_df, aes(x = Age,y = Prob, color = Mixture)) + 
    geom_line() + 
    labs(title = "Probability of Mixture by Age", x = "Age", y = "Probability of Being in each Mixture")+
    scale_color_viridis(discrete=TRUE,end = .95) 

ggplot(pi_df, aes(x = Age,y = Prob)) + 
    geom_line() + 
    labs(title = "Probability of Mixture by Age", x = "Age", y = "Probability of Being in each Mixture")+
    facet_wrap(vars(Mixture))
```

The two plots visualize the same data. In the first the probability of mixture by age for all mixtures is plotted together while the second plot has each mixture separated. 


# Transition

### Waking 

```{r}

tran_df$mixture <- as.factor(tran_df$mixture)
tran_df$weekend <- as.factor(tran_df$weekend)

tran_df_wake <-tran_df %>% filter(type == "Waking")

ggplot(tran_df_wake, aes(x = time,y = prob, color = mixture)) + 
    geom_line() + 
    labs(title = "Waking Probability", x = "Hour", y = "Probability of Waking") +
    facet_wrap(vars(weekend),labeller = labeller(
      weekend = c("1" = "Week", "2" = "Weekend")))+
    scale_color_viridis(discrete=TRUE,end = .95) 

ggplot(tran_df_wake, aes(x = time,y = prob, color = weekend)) + 
    geom_line() + 
    labs(title = "Waking Probability", x = "Hour", y = "Probability of Waking") +
    facet_wrap(vars(mixture))+
    scale_color_viridis(discrete=TRUE,end = .95,
                        name = "Day Type", labels = c("Week","Weekend")) 


```

The above 2 plots visualize the same data in two ways. In the first, wake probabilities for each mixture are plotted together and separated by week/weekend. In the second wake probability for week/weekend are plotted together and separated by mixture.

# Falling Asleep

```{r}
tran_df_sleep <-tran_df %>% filter(type != "Waking") 

ggplot(tran_df_sleep, aes(x = time,y = prob, color = mixture)) + 
    geom_line() + 
    labs(title = "Falling Asleep Probability", x = "Hour", y = "Probability of Falling Asleep") + 
    facet_wrap(vars(weekend),labeller = labeller(
      weekend = c("1" = "Week", "2" = "Weekend")))+
    scale_color_viridis(discrete=TRUE,end = .95) 

ggplot(tran_df_sleep, aes(x = time,y = prob, color = weekend)) + 
    geom_line() + 
    labs(title = "Falling Asleep Probability", x = "Hour", y = "Probability of Falling Asleep") + 
    facet_wrap(vars(mixture))+
    scale_color_viridis(discrete=TRUE,end = .95,
                        name = "Day Type", labels = c("Week","Weekend"))  


```

# Emission


```{r}


bi_dens_df<- data.frame(Activity = c(),
                               Light = c(),
                               MC = c(),
                        Density = c(),
                               Mixture = c(),
                               Week = c())

  

for (week_ind in 1:2){
  
  act_min <- min(emit_act[,1,,week_ind]) - 3*max(emit_act[,2,,week_ind])
  act_max <- max(emit_act[,1,,week_ind]) + 3*max(emit_act[,2,,week_ind])

  # if (week_ind == 2){act_max <-2}

  light_min <- min(emit_light[,1,,week_ind]) - 3*max(emit_light[,2,,week_ind])
  light_max <- max(emit_light[,1,,week_ind]) + 3*max(emit_light[,2,,week_ind])

  act_seq <- seq(act_min,act_max,length.out = 500)
  light_seq <- seq(light_min,light_max,length.out = 500)

  den_points <- cbind(rep(act_seq,length(light_seq)),
                            rep(light_seq,each = length(act_seq)))
  
  for (re_ind in 1:mix_num){
    for (mc_state in 1:2){
      
      emit_act_ind <- emit_act[mc_state,,re_ind,week_ind]
      emit_light_ind <- emit_light[mc_state,,re_ind,week_ind]
      
      mu <- c(emit_act_ind[1],emit_light_ind[1])
      sigma <- matrix(c(emit_act_ind[2]^2,emit_act_ind[2] * emit_light_ind[2] * corr_mat[re_ind,mc_state,week_ind],
                        emit_act_ind[2] * emit_light_ind[2] * corr_mat[re_ind,mc_state,week_ind],emit_light_ind[2]^2), ncol = 2)
      

      dens <- mvtnorm::dmvnorm(den_points,mu,sigma)
      dens <- dens/sum(dens)
      
      # sim_points <- mvtnorm::rmvnorm(10000,mu,sigma)
      # act_toremove <- quantile(sim_points[,1],.99)
      # light_toremove <- quantile(sim_points[,2],.99)
      # sim_points <- sim_points[(sim_points[,1] < act_toremove) & (sim_points[,2] < light_toremove),]
      
      bi_dens_df_working <- data.frame(Activity = den_points[,1],
                               Light = den_points[,2],
                               Density = dens,
                               MC = mc_state,
                               Mixture = re_ind,
                               Week = week_ind)
      
      bi_dens_df <- rbind(bi_dens_df,bi_dens_df_working)
      
    }
  }
}

      
bi_dens_df$MC <- as.factor(bi_dens_df$MC)
bi_dens_df$Mixture <- as.factor(bi_dens_df$Mixture)
bi_dens_df$Week <- as.factor(bi_dens_df$Week)



ggplot(bi_dens_df %>% filter(MC == 1), aes(x = Activity, y = Light, z = Density)) +
  geom_contour(aes(colour = after_stat(level)),bins = 500) + 
  scale_color_viridis(end = .95)+
  theme_bw() + 
  facet_grid(vars(Mixture),vars(Week),labeller = labeller(
      Week = c("1" = "Week", "2" = "Weekend")))+ 
  labs(x = "Activity", y = "Light", color = "Density",
       title = "Wake Activity and Light by Mixture and Week/Weekend") + 
  geom_hline(yintercept=-1.56,linetype="dashed") + 
  geom_vline(xintercept=-5.52,linetype="dashed")
```

This plot and the one below are density plots for the parameters of the estimated bivariate emissions distribution.

The dashed lines are the LoD for activity and light. This means that for any observations below the dashed horizontal line or to the left of the dashed vertical line we only observe the LoD. 


```{r}

ggplot(bi_dens_df %>% filter(MC == 2), aes(x = Activity, y = Light, z = Density)) +
  geom_contour(aes(colour = after_stat(level)),bins = 1000) + 
  scale_color_viridis(end = .95)+
  theme_bw() + 
  facet_grid(vars(Mixture),vars(Week),labeller = labeller(
      Week = c("1" = "Week", "2" = "Weekend")))+ 
  labs(x = "Activity", y = "Light", color = "Density",
       title = "Sleep Activity and Light by Mixture and Week/Weekend") + 
  geom_hline(yintercept=-1.56,linetype="dashed") + 
  geom_vline(xintercept=-5.52,linetype="dashed")
```


```{r, include=F}


Viterbi <- function(){
  decoded_array <- array(NA, dim = c(day_length,num_of_people,mix_num))
  for (ind in 1:num_of_people){
    for (clust_i in 1:mix_num){
      vit_ind_vec <- ViterbiIndHelper(ind,clust_i)
      decoded_array[,ind,clust_i] <- vit_ind_vec
    }
  }
  return(decoded_array)
}

ViterbiIndHelper <- function(ind,clust_i){
  
  
  tran_list_clust <- tran_list[[clust_i]]
  
  emit_act_week <- emit_act[,,,1]
  emit_act_weekend <- emit_act[,,,2]
  
  emit_light_week <- emit_light[,,,1]
  emit_light_weekend <- emit_light[,,,2]
  
  vcovar_vec <- vcovar_mat[,ind]
  
  
  log_class_0_week <- logClassificationC( act[,ind], light[,ind], 
                                          emit_act_week[1,1,clust_i], 
                                          emit_act_week[1,2,clust_i], 
                                          emit_light_week[1,1,clust_i],
                                          emit_light_week[1,2,clust_i], 
                                          lod_act, lod_light, corr_mat[clust_i,1,1], lintegral_mat[clust_i,1,1])
  
  log_class_1_week <- logClassificationC( act[,ind], light[,ind], 
                                          emit_act_week[2,1,clust_i], 
                                          emit_act_week[2,2,clust_i], 
                                          emit_light_week[2,1,clust_i],
                                          emit_light_week[2,2,clust_i], 
                                          lod_act, lod_light, corr_mat[clust_i,2,1], lintegral_mat[clust_i,2,1])  
  
  log_class_0_weekend <- logClassificationC( act[,ind], light[,ind], 
                                          emit_act_weekend[1,1,clust_i], 
                                          emit_act_weekend[1,2,clust_i], 
                                          emit_light_weekend[1,1,clust_i],
                                          emit_light_weekend[1,2,clust_i], 
                                          lod_act, lod_light, corr_mat[clust_i,1,2], lintegral_mat[clust_i,1,2])
  
  log_class_1_weekend <- logClassificationC( act[,ind], light[,ind], 
                                          emit_act_weekend[2,1,clust_i], 
                                          emit_act_weekend[2,2,clust_i], 
                                          emit_light_weekend[2,1,clust_i],
                                          emit_light_weekend[2,2,clust_i], 
                                          lod_act, lod_light, corr_mat[clust_i,2,2], lintegral_mat[clust_i,2,2])
  
  
  log_class_0 <- (log_class_0_week * (1-vcovar_vec)) + (log_class_0_weekend * vcovar_vec) 
  log_class_1 <- (log_class_1_week * (1-vcovar_vec)) + (log_class_1_weekend * vcovar_vec)
  
  
  viterbi_mat <- matrix(NA,2,day_length)
  viterbi_mat[1,1] <- log(init[clust_i,1]) + log_class_0[1]
  viterbi_mat[2,1] <- log(init[clust_i,2]) + log_class_1[1]
  
  viterbi_ind_mat <- matrix(NA,2,day_length)
  
  
  for (time in 2:day_length){
    
    tran <- tran_list_clust[[vcovar_vec[time]+1]][[time]]
    
    viterbi_mat[1,time] <- log_class_0[[time]] + 
      max(viterbi_mat[1,time-1] + log(tran[1,1]),
          viterbi_mat[2,time-1] + log(tran[2,1]))
    
    
    viterbi_mat[2,time] <- log_class_1[[time]] + 
      max(viterbi_mat[1,time-1] + log(tran[1,2]),
          viterbi_mat[2,time-1] + log(tran[2,2]))
    
    
    viterbi_ind_mat[1,time] <-  which.max(c(viterbi_mat[1,time-1] + log(tran[1,1]),
                                            viterbi_mat[2,time-1] + log(tran[2,1])))
    
    
    viterbi_ind_mat[2,time] <- which.max(c(viterbi_mat[1,time-1] + log(tran[1,2]),
                                           viterbi_mat[2,time-1] + log(tran[2,2])))
    
    
  }
  
  decoded_mc <- c(which.max(viterbi_mat[,time]))
  for(time in day_length:2){
    decoded_mc <- c(viterbi_ind_mat[decoded_mc[1],time],decoded_mc)
  }
  
  return(decoded_mc-1)
} 

GenTranList <- function(params_tran_array,time_vec,mix_num,vcovar_num){
  
  
  mixture_vcovar_tran_list <- list()
  vcovar_tran_list <- list()
  
  for (mixture_ind in 1:mix_num){
    for (vcovar_ind in 1:vcovar_num){
      
      vcovar_tran_list[[vcovar_ind]] <- TranByTimeVec(re_ind = mixture_ind,
                                                      params_tran = params_tran_array[,,vcovar_ind],
                                                      time_vec = time_vec)
      
    }
    mixture_vcovar_tran_list[[mixture_ind]] <- vcovar_tran_list
  }
  
  return(mixture_vcovar_tran_list)
}

TranByTimeVec <- function(re_ind, params_tran,time_vec){
  return(lapply(time_vec, Params2Tran, params_tran = params_tran,re_ind=re_ind))
}

readCpp <- function(path) {
  tryCatch(
    {
      sourceCpp(file = path)
    },
    error = function(cond) {
      message("Wrong environment")
      # Choose a return value in case of error
      NA
    },
    warning = function(cond) {
      message("Wrong environment")
      # Choose a return value in case of warning
      NULL
    },
    finally = {
      message("Done")
    }
  )
}

CalcLintegralMat <- function(emit_act,emit_light,corr_mat,lod_act,lod_light){
  mix_num <- dim(emit_act)[3]
  if (is.na(mix_num)){mix_num <- 1}
  
  lintegral_mat <- array(NA,dim = c(mix_num,2,2))
  #j is week/weekend
  for (j in 1:2){
    for (i in 1:mix_num){
      lintegral_mat[i,1,j] <- log(integrate(Case4,lower = -Inf,upper = lod_act,
                                          emit_act[1,1,i,j],emit_act[1,2,i,j],
                                          emit_light[1,1,i,j],emit_light[1,2,i,j],
                                          corr_mat[i,1,j],lod_light)[[1]])
      
      lintegral_mat[i,2,j] <- log(integrate(Case4,lower = -Inf,upper = lod_act,
                                          emit_act[2,1,i,j],emit_act[2,2,i,j],
                                          emit_light[2,1,i,j],emit_light[2,2,i,j],
                                          corr_mat[i,2,j],lod_light)[[1]])
    }
  }
  
  lintegral_mat[lintegral_mat == -Inf] <- -9999
  
  return(lintegral_mat)
}

Case4 <- function(act_obs,mu_act,sig_act,mu_light,sig_light,bivar_corr,light_LOD){
  
  mu_light_cond <- CalcCondMean(mu_light,sig_light,mu_act,sig_act,bivar_corr,act_obs)
  sig_light_cond <- CalcCondSig(sig_light,bivar_corr)
  
  lognorm_dens <- dnorm(act_obs,mu_act,sig_act) * 
    pnorm(light_LOD,mu_light_cond,sig_light_cond)
  return(lognorm_dens)
}

#Calculates mu1|2
CalcCondMean <- function(mu1,sig1,mu2,sig2,bivar_corr,obs2){
  return(mu1 + bivar_corr*(sig1/sig2)*(obs2-mu2))
}

#Calculates sig1|2
CalcCondSig <- function(sig1,bivar_corr){
  return(sig1*sqrt(1-bivar_corr^2))
}
```

```{r,include=F}

lintegral_mat <- CalcLintegralMat(emit_act,emit_light,corr_mat,lod_act,lod_light)


readCpp( "../cFunctions.cpp" )
tran_list <- GenTranList(params_tran_array,c(1:day_length),mix_num,2)

decoded_mat <- Viterbi()


DecodingPlots <- function(ind,act_bool){
  ind_age <- age_vec[ind]
  
  time_ticks <- 28
  
  
  sub = paste("JMHMM with", mix_num, "Mixtures")
  time_list <- seq(as.POSIXct("2024-01-01"), as.POSIXct("2024-01-10"), by=60*15)
  time_list <- time_list[-length(time_list)]
  
  decoded_df <- data.frame(decode = as.vector(decoded_mat[,ind,mix_assignment[ind]]),
                           act = act[,ind],
                           light = light[,ind],
                           time =time_list)
  
  
  if (act_bool){
    max_dec <- round(max(act[,ind],na.rm = T))
    min_dec <- round(min(act[,ind],na.rm = T))
    obs_type <- "Activity"
  } else {
    max_dec <- round(max(light[,ind],na.rm = T))
    min_dec <- round(min(light[,ind],na.rm = T))
    obs_type <- "Light"
  }
  
  decoded_df <- decoded_df %>% 
    mutate(decode = case_when(decode == 0 ~ max_dec,
                              decode == 1 ~ min_dec))
  
  if (act_bool){
    g <- ggplot(decoded_df, aes(x=time, y=act))
  } else {
    g <- ggplot(decoded_df, aes(x=time, y=light))
  }
  
  g <- g +
    geom_path(color = viridis(20)[9],aes(x=time, y=decode)) +
    geom_point(aes(col = as.factor(decode)),size = 1) +
    scale_x_datetime(breaks= seq(min(decoded_df$time), max(decoded_df$time)+15*60, length=time_ticks), 
                     date_labels="%H:%M")+
    theme_bw() +
    labs(x = "Time",
         y = "Predicted State",
         title = paste("Predicted State and Observed",obs_type,"by Time for Person", ind),
         subtitle = sub) +
    scale_y_continuous(breaks=c(min_dec,max_dec),
                       labels = c("Sleep","Wake"),
                       sec.axis = sec_axis(~ . * 1,
                                           name = "Acitivity (Log MIMS)",
                                           breaks=seq(min_dec,max_dec,2)))+
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          axis.ticks.y.left = element_line(color = c("black", "black")),
          legend.position="bottom") +
    scale_colour_viridis_d(end = .9,name = "Predicted State", labels = c("Sleep", "Wake"))
  
 
  return(g)
}


which(mix_assignment == 8)

DecodingPlots(10,T)
DecodingPlots(10,F)
```

```{r}
ReturnLongestBlockIndex <- function(vit_ind_day,forw_look){
  if (sum(vit_ind_day) == 0){return(0)}
  
  if (sum(vit_ind_day[1:(96-forw_look)])==0){
    return(c(which.max(vit_ind_day),96))
  }
  
  forw_look <- forw_look - 1
  rs <- numeric((length(vit_ind_day)-forw_look))
  for (i in 1:(length(vit_ind_day)-forw_look)){
    if (i > 1){
      rs[i] <- sum(vit_ind_day[i:(i+forw_look)])*max(rs[i-1]>0,vit_ind_day[i])
    } else {
      rs[i] <- sum(vit_ind_day[i:(i+forw_look)])*vit_ind_day[i]
    }
  }
  
  rs <- rs > 0
  runs <- rle(rs)
  lengths <- runs$lengths
  
  end = cumsum(lengths)
  start = c(1, lag(end)[-1] + 1)
  
  is_max <- which(lengths == max(lengths[runs$values]) & runs$values)
  return(c(start[is_max],end[is_max]))
}


full_data_ind <- which(colSums(!is.na(act)) > 96*6 & colSums(!is.na(light)) > 96*6)
num_of_days <- 8
viterbi_df <- data.frame()

ndf_ent <- length(full_data_ind) * mix_num * 3

keep_time <- c()
keep_array <- array(NA, dim = dim(decoded_mat))

viterbi_df <- data.frame(SleepBlock = numeric(ndf_ent),
                         SleepEff = numeric(ndf_ent),
                         SleepOnset = numeric(ndf_ent),
                         SleepTotal = numeric(ndf_ent),
                         Mixture = numeric(ndf_ent),
                         WeekendInd = numeric(ndf_ent),
                         Weight = numeric(ndf_ent),
                         Ind = numeric(ndf_ent),
                         pov_cont = numeric(ndf_ent),
                         age_cont = numeric(ndf_ent),
                         race = numeric(ndf_ent),
                         bmi_cont = numeric(ndf_ent),
                         bmi_disc = numeric(ndf_ent),
                         gender = numeric(ndf_ent),
                         houseFS = numeric(ndf_ent),
                         insurance = numeric(ndf_ent),
                         income = numeric(ndf_ent),
                         depression = numeric(ndf_ent),
                         home_own = numeric(ndf_ent))
row_track <- 1

for (ind in full_data_ind){
  
  demo_vars <-c(ind,
                id$poverty[ind],
                id$age[ind],
                id$race[ind],
                id$BMI[ind],
                id$bmi_disc[ind],
                id$gender[ind],
                id$houseFS[ind],
                id$insurance[ind],
                id$income[ind],
                id$depression[ind],
                id$home_own[ind])
  
  
  vcovar_vec <- vcovar_mat[,ind]
  
  for(clust_i in 1:mix_num){
    keep_time_ind <- c()
    working_vit_df <- data.frame()
    vit_ind <- decoded_mat[,ind,clust_i]
    
    for(day in 1:num_of_days){
      
      start_ind <- 69 + (96*(day-1))
      stop_ind <- 68 + (96*day)
      vit_ind_day <- vit_ind[start_ind:stop_ind]
      
      vcovar_ind_day <- vcovar_vec[start_ind:stop_ind]
      
      block_indices <- ReturnLongestBlockIndex(vit_ind_day,2)
      
      if (length(block_indices)==1){
        sleep_block <- 0
        sleep_eff <- 0
        sleep_onset <- 0
        sleep_total <- sum(vit_ind_day == 1)
        keep_time_ind <- c(keep_time_ind,c(1:96) %in% 0)
      
      } else {
        block_start <- block_indices[1]
        block_end <- block_indices[2]
        
        sleep_block <- block_end - block_start + 1
        sleep_eff <- sum(vit_ind_day[block_start:block_end]==1)/sleep_block
        sleep_onset <- block_start + 48
        sleep_total <- sum(vit_ind_day == 1)
        
        keep_time_ind <- c(keep_time_ind,c(1:96) %in% c(block_start:block_end))
      }
    
    # This used to be 32 not 48
      if (sum(is.na(act[start_ind:stop_ind,ind]))>48){
        sleep_block <- NA
        sleep_eff <- NA
        sleep_onset <- NA
        sleep_total <- NA
      }
      
      if (all(vcovar_ind_day== 0) ){weekend_ind <- 0
      } else if (all(vcovar_ind_day== 1) ){weekend_ind <- 1
      } else {print("Mismatch time and weekend")}
      
      working_vit_df <- rbind(working_vit_df,
                              c(sleep_block,sleep_eff,sleep_onset,sleep_total,
                                clust_i,weekend_ind,re_prob[ind,clust_i]))
      
    }
    colnames(working_vit_df) <- c("Block","Eff","Onset","Total","Mixture","WeekendInd","Weight")
    
    total_vals <- apply(working_vit_df,2,median, na.rm = T)
    #change weekend vals to 3 for total
    total_vals[6] <- 2
    
    week_vals <- apply(working_vit_df %>% filter(WeekendInd==0),2,median, na.rm = T)
    weekend_vals <- apply(working_vit_df %>% filter(WeekendInd==1),2,median, na.rm = T)
    
    
    viterbi_df[row_track,] <- c(total_vals,demo_vars)
    viterbi_df[row_track+1,] <- c(week_vals,demo_vars)
    viterbi_df[row_track+2,] <- c(weekend_vals,demo_vars)
    row_track <- row_track + 3
    
    keep_time_ind <- c(rep(F,68),keep_time_ind,rep(F,28))
    keep_array[,ind,clust_i] <- keep_time_ind
    
  
  }
}



viterbi_df$WeekendInd <- as.factor(viterbi_df$WeekendInd)
viterbi_df$Mixture <- as.factor(viterbi_df$Mixture)
viterbi_df$race <- as.factor(viterbi_df$race)
# 
viterbi_df$gender <- as.factor(viterbi_df$gender)
viterbi_df$houseFS <- as.factor(viterbi_df$houseFS)


viterbi_df$insurance <- as.factor(viterbi_df$insurance)
viterbi_df$income <- as.factor(viterbi_df$income)
viterbi_df$home_own <- as.factor(viterbi_df$home_own)


```

```{r}

ggplot(viterbi_df, aes(x=Mixture, y=SleepBlock, fill=WeekendInd,weight = Weight)) +
  geom_boxplot(position=position_dodge(.8))


ggplot(viterbi_df, aes(x=Mixture, y=SleepOnset, fill=WeekendInd,weight = Weight)) +
  geom_boxplot(position=position_dodge(.8))


ggplot(viterbi_df, aes(x=Mixture, y=SleepEff, fill=WeekendInd,weight = Weight)) +
  geom_boxplot(position=position_dodge(.8))


ggplot(viterbi_df, aes(x=Mixture, y=SleepTotal, fill=WeekendInd,weight = Weight)) +
  geom_boxplot(position=position_dodge(.8))





ggplot(viterbi_df, aes(x=Mixture, y=pov_cont,weight = Weight)) +
  geom_boxplot()
ggplot(viterbi_df, aes(x=Mixture, y=bmi_cont,weight = Weight)) +
  geom_boxplot()
ggplot(viterbi_df, aes(x=Mixture, y=age_cont,weight = Weight)) +
  geom_boxplot()
ggplot(viterbi_df, aes(x=Mixture, y=depression,weight = Weight)) +
  geom_boxplot()


ggplot(viterbi_df, aes(fill=gender, y=Weight, x=Mixture)) +
    geom_bar(position="fill", stat="identity")

ggplot(viterbi_df, aes(fill=houseFS, y=Weight, x=Mixture)) +
    geom_bar(position="fill", stat="identity")

ggplot(viterbi_df, aes(fill=race, y=Weight, x=Mixture)) +
    geom_bar(position="fill", stat="identity")


ggplot(viterbi_df, aes(fill=insurance, y=Weight, x=Mixture)) +
    geom_bar(position="fill", stat="identity")

ggplot(viterbi_df, aes(fill=income, y=Weight, x=Mixture)) +
    geom_bar(position="fill", stat="identity")

ggplot(viterbi_df, aes(fill=home_own, y=Weight, x=Mixture)) +
    geom_bar(position="fill", stat="identity")


```

```{r}
library(pec)
library(systemfonts)
library(StepReg)

surv_data2 <- viterbi_df %>% filter(WeekendInd == 2)
surv_data2 <- surv_data2 %>% mutate(time = rep(surv_data$time[full_data_ind],each = mix_num)) %>%
  mutate(status = rep(surv_data$status[full_data_ind],each = mix_num))
surv_data2 <- surv_data2 %>% filter(Weight > 0)

surv_data2 <- surv_data2 %>% select (-c(bmi_disc))

surv_data2$race <- as.factor(surv_data2$race)
surv_data2$gender <- as.factor(surv_data2$gender)
surv_data2$houseFS <- as.factor(surv_data2$houseFS)
surv_data2$Mixture <- as.factor(surv_data2$Mixture)
surv_data2$insurance <- as.factor(surv_data2$insurance)
surv_data2$income <- as.factor(surv_data2$income)
surv_data2$home_own <- as.factor(surv_data2$home_own)

var_to_inc <- colnames(surv_data2)
var_to_inc <- var_to_inc[!(var_to_inc %in% c("time","status","Weight"))]

fullform <- as.formula(paste("Surv(time,status) ~ ",paste(var_to_inc,collapse = " + ")))
# 
# selected_var <- selectCox(fullform, data = surv_data2,rule = "aic")$In
# selectform <- as.formula(paste("Surv(time,status) ~ ",paste(selected_var,collapse = " + ")))
# coxph(selectform, data = surv_data2)
# 
# 
# backward2 <- stepwise(formula = fullform,
#                 data = surv_data2,
#                 type = "cox",
#                 strategy = "backward",
#                 metric = c("AICc"))
# 
# 
# bidirection <- stepwise(formula = fullform,
#                 data = surv_data2,
#                 type = "cox",
#                 strategy = "bidirection",
#                 metric = c("AICc"),
#                 sle = .05,
#                 sls = .05)
# 
# 
# plot(bidirection)
```

```{r}
fit <- coxph(Surv(time, status) ~ age_cont + Mixture + pov_cont + race + gender + bmi_cont + home_own + SleepEff + SleepBlock + SleepOnset+ SleepTotal,weights = Weight, data = surv_data2)
fit

fit <- coxph(Surv(time, status) ~ age_cont + Mixture + depression + pov_cont + race + gender + bmi_cont + home_own + SleepEff + SleepBlock + SleepOnset+ SleepTotal,weights = Weight, data = surv_data2)
fit


fit <- coxph(Surv(time, status) ~ age_cont + Mixture,weights = Weight, data = surv_data2)
fit


# (testfit <- cox.zph(fit))
# plot(testfit)
# 
# event_ind <- which(surv_data2$status == 1)
# nevent_ind <- which(surv_data2$status != 1)
# 
# mresid <- (surv_data2$status) - predict(fit, type='expected')
# 
# plot(mresid[event_ind])
# plot(mresid[nevent_ind])
```

```{r}
library(matrixStats)
CalcBLHaz <- function(beta_age,vec, re_prob,surv_event,surv_time,age_vec){
  n <- length(surv_event)
  bline_vec <- numeric(n)
  cbline_vec <- numeric(n)
  
  for (time_ind in 1:n){
    risk_set <- surv_time >= surv_time[time_ind]
    denom <- sum((re_prob[risk_set,] %*% exp(beta_vec)) * exp(beta_age * age_vec[risk_set]))
    bline_vec[time_ind] <- surv_event[time_ind]/denom
  }
  
  for(time_ind in 1:n){
    anti_risk_set <- surv_time <= surv_time[time_ind]
    cbline_vec[time_ind] <- sum(bline_vec[anti_risk_set])
  }
  
  return(list(bline_vec,cbline_vec))
}

CalcProbRE <- function(alpha,pi_l){
  
  len <- dim(alpha[[1]])[1]
  re_len <- dim(alpha[[1]])[3]
  re_weight_vec <- numeric(re_len)
  re_weights <- matrix(0,nrow = length(alpha),ncol = re_len)
  
  
  for (ind in 1:length(alpha)){
    for (re_ind in 1:re_len){
      re_weights[ind,re_ind] <- logSumExp(alpha[[ind]][len,,re_ind]) + log(pi_l[ind,re_ind])
    }
    re_weights[ind,] <- exp(re_weights[ind,] - logSumExp(c(re_weights[ind,])))
    
  }
  
  return(re_weights)
  
}

CollapseVitKeepArrays <- function(decoded_mat,mix_assignment){
  flat_decoded_mat <- matrix(0,dim(decoded_mat)[1],dim(decoded_mat)[2])
  flat_keep_mat <- matrix(0,dim(decoded_mat)[1],dim(decoded_mat)[2])
  for (ind in 1:num_of_people){
    flat_decoded_mat[,ind] <- decoded_mat[,ind,mix_assignment[ind]]
    flat_keep_mat[,ind] <- keep_array[,ind,mix_assignment[ind]]
  }
  return(list(flat_decoded_mat,flat_keep_mat))
}

flat_array_list <- CollapseVitKeepArrays(decoded_mat,mix_assignment)
flat_decoded_mat <- flat_array_list[[1]]
flat_keep_mat <- flat_array_list[[2]]

bhaz_vec <- CalcBLHaz(beta_age,beta_vec,re_prob,surv_event,surv_time,age_vec)
bline_vec <- bhaz_vec[[1]]
cbline_vec <- bhaz_vec[[2]]



flat_decoded_mat_full <- flat_decoded_mat[,full_data_ind]
flat_decoded_mat_full[!flat_keep_mat[,full_data_ind]] <- NA

library(Rcpp)
readCpp( "~/Documents/Predoc/JointHMM/cFunctions.cpp" )
  

# alpha_full <- ForwardAltC(flat_decoded_mat_full,init,tran_list,beta_vec,beta_age,
#                      surv_event[full_data_ind],bline_vec[full_data_ind],
#                      cbline_vec[full_data_ind],age_vec[full_data_ind],vcovar_mat[,full_data_ind])
# 
# re_prob_recon <- CalcProbRE(alpha_full,pi_l[full_data_ind,])
# 
# tab <- table(apply(re_prob_recon,1,which.max),mix_assignment[full_data_ind])
# 
# tab / replicate(mix_num,rowSums(tab))


alpha <- ForwardAltC(flat_decoded_mat,init,tran_list,beta_vec,beta_age,
                     surv_event,bline_vec,
                     cbline_vec,age_vec,vcovar_mat)

re_prob_recon <- CalcProbRE(alpha,pi_l)

tab <- table(apply(re_prob_recon,1,which.max),mix_assignment)

#Sensitivity
round(tab / t(replicate(mix_num,colSums(tab))),4)

#Specificity
round(tab / replicate(mix_num,rowSums(tab)),4)


pred_vec <- numeric(num_of_people)
for (i in 1:num_of_people){pred_vec[i] <- re_prob_recon[i,mix_assignment[i]]}

hist(pred_vec[mix_assignment==1])
hist(pred_vec[mix_assignment==2])
hist(pred_vec[mix_assignment==3])
```
